\documentclass[10pt,a4paper,oneside,notitlepage]{report}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage[table]{xcolor}  
\usepackage[margin=2.5cm]{geometry}
\usepackage{titling}
\usepackage{titlesec}
\usepackage{graphicx}
\usepackage{float}

% kein einrücken bei neuen absätzen
\setlength{\parindent}{0pt}

% position titel
\setlength{\droptitle}{-2cm}

% abstand vor subsection titel
\titlespacing{\subsection}{0pt}{40pt}{8pt}

\author{Flurin Rindisbacher}
\title{How To Write Fast Numerical Code \\ \vspace{6 mm} \textbf{Assignment 3}}

\begin{document}
\maketitle

\section*{Exercise 1 - Cache mechanics}
\subsection*{(a)}
There are $32KB / 64 Byte = 512$ cache blocks. The number of sets is therefore $\underline{512 / 8=64}$.

\subsection*{(b)}
The block size is $64$ bytes and with $64=2^6$ 6 bits are needed to determine the position in the block. \\
Also 6 bits are needed for the set. \\
\underline{This results in a tag length of $64 - 6 - 6=52$ bits.}
\\
\textbf{TODO:} effectly used on this machine??

\subsection*{(c)}
$0xDE147BA$ results ins the following values: \\
tag: $0xDE14$ \\
set: $0x1E$ \\
block: $0x3A$

\section*{Exercise 2 - Cache mechanics}
cache size: 32KB \\
block size: 64 bytes \\
number of sets: 32KB / 16bytes = 1024 sets \\
because of sizeof(int) = 4, 4 integers fit into one block. \\
address of src matrix: 0x00 \\
address of dest matrix assuming size 64x64: 0x4000 ($64*64*4=16384$) \\
address of dest matrix assuming size 96x64 = 0x6000 ($96*64*4=25576$) \\
Number of bits for block: 4 \\
Number of bits for set: 10\\
\subsection*{(a)}
\subsubsection*{i}
with 4 integers per block and 64 ints per row there are 16 blocks to read/write per row. 0x00 (src) and 0x4000 (dest) map to the same cache set. we therefore know, that they'll use the same 16 blocks for the dest[i] and src[i] row.  \\ Assuming the used blocks are numbered from 0 to 15 the access pattern will be the following: \\

\begin{tabular}{|c|c|c|c|}
\hline 
\rowcolor{gray!30}
\textbf{block for src} &\textbf{ block for dest} & \textbf{hit/miss src} & \textbf{hit/miss dest} \\ 
\hline 
0 & 15 & MHHH & MHHH \\ 
\hline 
1 & 14 & MHHH & MHHH \\ 
\hline 
2 & 13 & MHHH & MHHH \\ 
\hline 
3 & 12 & MHHH & MHHH \\ 
\hline 
4 & 11 & MHHH & MHHH \\ 
\hline 
5 & 10 & MHHH & MHHH \\ 
\hline 
6 & 9 & MHHH & MHHH \\ 
\hline 
7 & 8 & MHHH & MHHH \\ 
\hline 
8 & 7 & MHHH & MHHH \\ 
\hline 
9 & 6 & MHHH & MHHH \\ 
\hline 
10 & 5 & MHHH & MHHH \\ 
\hline 
11 & 4 & MHHH & MHHH \\ 
\hline 
12 & 3 & MHHH & MHHH \\ 
\hline 
13 & 2 & MHHH & MHHH \\ 
\hline 
14 & 1 & MHHH & MHHH \\ 
\hline 
15 & 0 & MHHH & MHHH \\ 
\hline 
\end{tabular} 

Because the src / dest never uses the same block at the same time we always get an access pattern of MHHH. This results in a cache miss rate of $\underline{1/4=0.25}$

\subsubsection*{ii}
0x00 (src) points to set number 0. 0x6000 (dest) points to set number 512.
So when src reads the cache block $b=(i \mod 1024)$ dest uses block $b=(i+512 \mod 1024)$. This always gives a MHHH access pattern for read and write. In total thats a cache miss rate of $\underline{1/4=0.25}$

\end{document}
